Computing in the Field of Rationals

### 2012-06-21 rational_ext.tex
_

The operational semantics of {{Pi}} is summarized below (also see
\cite{infeffects}).  The semantics of the primitive combinators is
given by the following single-step reductions below. Since there are
no values of type {{0}}, the rules omit the impossible cases:
\begin{scriptsize}
%subcode{opsem} include main
%! columnStyle = rlcl
% swap+ & (left v) &|-->& right v
% swap+ & (right v) &|-->& left v 
% assocl+ & (left v1) &|-->& left (left v1)
% assocl+ & (right (left v2)) &|-->& left (right v2)
% assocl+ & (right (right v3)) &|-->& right v3 
% assocr+ & (left (left v1)) &|-->& left v1
% assocr+ & (left (right v2)) &|-->& right (left v2)
% assocr+ & (right v3) &|-->& right (right v3)
% identl* & ((), v) &|-->& v 
% identr* & v &|-->& ((), v) 
% swap* & (v1, v2) &|-->& (v2, v1) 
% assocl* & (v1, (v2, v3)) &|-->& ((v1, v2), v3) 
% assocr* & ((v1, v2), v3) &|-->& (v1, (v2, v3)) 
% dist & (left v1, v3) &|-->& left (v1, v3)
% dist & (right v2, v3) &|-->& right (v2, v3)
% factor & (left (v1, v3)) &|-->& (left v1, v3) 
% factor & (right (v2, v3)) &|-->& (right v2, v3)

The operational semantics of the closure conditions are presented in
the usual big-step style.

%subcode{proof} include main
%@ ~
%@@ id v |--> v 
%
%@ c{dagger} v1 |--> v2
%@@ (sym c) v1 |--> v2
%
%@ c1  v1 |--> v
%@ c2  v |--> v2
%@@ (c1(;)c2)  v1 |--> v2
%---
%@ c1  v1 |--> v2
%@@ (c1 (+) c2)  (left v1) |--> left v2
%
%@ c2  v1 |--> v2
%@@ (c1 (+) c2)  (right v1) |--> right v2
%---
%@ c1  v1 |--> v3
%@ c2  v2 |--> v4
%@@ (c1 (*) c2)  (v1, v2) |--> (v3, v4)
\end{scriptsize}

@@

  \begin{proof}
    To establish that category is monoidal one must show isomorphisms
    \begin{enumerate}
    \item {{b1 + (b2 + b3) <-> (b1 +b2) + b3}} given by {{assoc+}}
    \item {{0 + b <-> b}} given by {{zeroe}}
    \item {{b + 0 <-> b}} given by {{swap+ (;) zeroe}}
    \item We need to check that {{+}} is a bifunctor. 
      \begin{enumerate}
      \item {{id_{b1}(+)id_{b2} = id_{b1+b2} }}
      \item {{(f(+)g)(;)(j(+)k) = (f (;) j) (+) (g (;) k)}}
      \end{enumerate}
    \item We need to check the naturality of {{assoc+}}, {{zeroe}} and
      {{swap+(;)zeroe}}.
      \begin{enumerate}
      \item {{assoc+ (;) ((f (+)g) (+) h) = (f (+)(g (+) h)) (;) assoc+}}
      \item {{zeroe (;) f = (id (+) f) (;) zeroe}}
      \item {{(swap+ (;) zeroe) (;) f = (f (+) id) (;) (swap+ (;) zeroe)}}
      \end{enumerate}
    \item Satisfy certain coherence conditions which are usually
      called the pentagon and traingle axioms.
    \end{enumerate}

>>>>

The operational semantics of {{Pi}} is summarized below (also see
\cite{infeffects}).  The semantics of the primitive combinators is
given by the following single-step reductions below. Since there are
no values of type {{0}}, the rules omit the impossible cases:
\begin{scriptsize}
%subcode{opsem} include main
%! columnStyle = rlcl
% swap+ & (left v) &|-->& right v
% swap+ & (right v) &|-->& left v 
% assocl+ & (left v1) &|-->& left (left v1)
% assocl+ & (right (left v2)) &|-->& left (right v2)
% assocl+ & (right (right v3)) &|-->& right v3 
% assocr+ & (left (left v1)) &|-->& left v1
% assocr+ & (left (right v2)) &|-->& right (left v2)
% assocr+ & (right v3) &|-->& right (right v3)
% identl* & ((), v) &|-->& v 
% identr* & v &|-->& ((), v) 
% swap* & (v1, v2) &|-->& (v2, v1) 
% assocl* & (v1, (v2, v3)) &|-->& ((v1, v2), v3) 
% assocr* & ((v1, v2), v3) &|-->& (v1, (v2, v3)) 
% dist & (left v1, v3) &|-->& left (v1, v3)
% dist & (right v2, v3) &|-->& right (v2, v3)
% factor & (left (v1, v3)) &|-->& (left v1, v3) 
% factor & (right (v2, v3)) &|-->& (right v2, v3)   
\end{scriptsize}
The operational semantics of the closure conditions are presented in
the usual big-step style.
\begin{scriptsize}
%subcode{proof} include main
%@ ~
%@@ id v |--> v 
%
%@ c{dagger} v1 |--> v2
%@@ (sym c) v1 |--> v2
%
%@ c1  v1 |--> v
%@ c2  v |--> v2
%@@ (c1(;)c2)  v1 |--> v2
%---
%@ c1  v1 |--> v2
%@@ (c1 (+) c2)  (left v1) |--> left v2
%
%@ c2  v1 |--> v2
%@@ (c1 (+) c2)  (right v1) |--> right v2
%---
%@ c1  v1 |--> v3
%@ c2  v2 |--> v4
%@@ (c1 (*) c2)  (v1, v2) |--> (v3, v4)
\end{scriptsize}

@@

  \begin{proof}
    To establish that category is monoidal one must show isomorphisms
    \begin{enumerate}
    \item {{b1 + (b2 + b3) <-> (b1 +b2) + b3}} given by {{assoc+}}
    \item {{0 + b <-> b}} given by {{zeroe}}
    \item {{b + 0 <-> b}} given by {{swap+ (;) zeroe}}
    \item We need to check that {{+}} is a bifunctor. 
      \begin{enumerate}
      \item {{id_{b1<->b1}(+)id_{b2<->b2} = id_{b1+b2<->b1+b2} }}
      \item {{(f(+)g)(;)(j(+)k) = (f (;) j) (+) (g (;) k)}}
      \end{enumerate}
    \item We need to check the naturality of {{assoc+}}, {{zeroe}} and
      {{swap+(;)zeroe}}.
      \begin{enumerate}
      \item {{assoc+ (;) ((f (+)g) (+) h) = (f (+)(g (+) h)) (;) assoc+}}
      \item {{zeroe (;) f = (id (+) f) (;) zeroe}}
      \item {{(swap+ (;) zeroe) (;) f = (f (+) id) (;) (swap+ (;) zeroe)}}
      \end{enumerate}
    \item Satisfy certain coherence conditions which are usually
      called the ``pentagon'' and ``triangle'' axioms.
    \end{enumerate}
